Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-01-31 paper/SupMat.Rnw
commit of ongoing edits in order to do a pull

%header material for paper
\title{Asymmetric relationships and their impact on coexistence: Supporting information}
\date{}

@@

%***DAN: I decided to present the general theory (not for the lottery model but for the
%general case) in the Theory section in the main text, putting off the specific lottery
%model application to later. For that, I need to explain scaling factors in general.
%So this section is now for that explanation. That means the below text, which is about
%the scaling factors for the lottery model need to be handled later. It is good, however,
%that the below text is for the general lottery model, for either way of doing the 
%fecundities. That will come in handy later.
%
% For the lottery model, for $i \neq j$, we have
% \begin{align}
% \sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
% \sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
% \sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
% \end{align}
% Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
% $B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
% in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
% but we instead compute $q_{ij}$ through implicit differentiation:
% \begin{align}
% \frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
% &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
% \end{align}
% Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
% \begin{align}
% q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
% &= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
% \end{align}
% as stated in the main text.

@@

We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
original decompositions of \cite{Chesson_1994}; and then we include our extensions of the decomposition.

@@

%***DAN: stopped here on 2023 01 27. See outline in the main text of how I need to restructure the 
%SI sections from this point forward.

@@

For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$, which is an increasing
function of $E_i$ and decreasing function of $C_i$.

When $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
Since we assumed log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SIsect:scaling_factors}.
Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.

\section{Mathematical simplifications for plotting}\label{sect:math_simplifications}

We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the lottery model. 

@@

E_i=B_i=\exp(b_i)

@@

C_i=\frac{B_1 N_1 + B_2 N_2}{\delta N}
C_{i \bs i} = B_j/\delta=\exp(b_j)/\delta
r_i=\ln(1-\delta+E_i/C_i)

@@

\olr_{i \bs i} &= \E[\ln(1-\delta+E_i/C_{i \bs i})] \\
&= \E[\ln(1-\delta+\delta B_i/B_j)] \\
&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]\label{eq:base_IGR}

@@

\olep_i^E &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i+\mu_i)/\exp(\mu_j+\sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}

@@

where $\tilde{b}_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(b_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma \tilde{b}_j +\mu_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(-\sigma \tilde{b}_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 

@@

@@

\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(b_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i +\mu_i)/\exp(\sigma \tilde{b}_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}

@@

where $(\tilde{b}_i, \tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$. Continuing,

@@

\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(b_i^\shp)/\exp(b_j^\shp))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i^\shp +\mu_i)/\exp(\sigma \tilde{b}_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}

@@

where $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,

@@

\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}

@@

where $(\tilde{b}_i,\tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$, and 
$\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normal and independent. Continuing,

@@

\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}

@@

where $(\ddot{b}_i,\ddot{b}_j)$ have the same (normal) marginals and covariance as $(b_i,b_j)$ but are 
bivariate normal. Thus
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})

@@

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}

@@

This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.

>>>>

%header material for paper
\title{Asymmetric relationships and their effects on coexistence: Supporting information}
\date{}

@@

@@

We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
original decompositions of \cite{Chesson_1994}; and then we include our extension of the decomposition.

@@

@@

\subsection{Theory that applies to both versions of the model}\label{SIsect:TheoryForLotteryBoth}

Recall we took $E_i(t)=B_i(t)$, thereby assuming that fecundity depends strictly on
the environment for the lottery model (both the log-normal fecundities and the beta fecundities versions of the model). 
Recall also that competition, $C_i(t)$, is 
$$C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N)$$ (again for both versions of the model), which
does not depend on $i$. Species growth rates for the lottery model, $r_i(t)=\ln(N_{i}(t+1)/N_i(t))$, can then be written 
$$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)],$$ 
which is an increasing
function of $E_i$ and decreasing function of $C_i$, as was assumed at the outset of Theory.

When $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$. Thus 
$$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta.$$
Lottery-model values of all the expressions of SI section \ref{SIsect:other_mechanisms} can be obtained 
from the above expression for $C_{i \bs i}$ and the assumption that $E_i(t)=B_i(t)$.
For instance, 
$$\olr_{i \bs i} = \E[r_i(E_i,C_{i \bs i})] = \E[\ln(1-\delta+E_i /C_{i \bs i})] = \E[\ln(1-\delta+\delta B_i/B_j)].$$ 
Likewise, 
$$\olr_{j \bs i} = \E[r_j(E_j,C_{j \bs i})] = \E[\ln(1-\delta+E_j/C_{j \bs i})] = \E[\ln(1-\delta+\delta B_j/B_j)]=0,$$
though we already knew, \emph{a priori}, that this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus 
$$\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)].$$ 
Likewise, $\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we can use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)].$$
An expression for $\olrn_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but with ATAs removed, $(B_i^\nat,B_j^\nat /\delta)$.
Thus 
$$\olrn_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\nat/B_j^\nat)].$$ 
Likewise, $\olrn_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})=(B_j,B_j/\delta)$ in an analogous way.
But note that $B_j$ and $B_j/\delta$ are linearly related for the lottery model, so their association is 
already symmetric, so
so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
This type of reasoning is applied to the log-normal fecundities and beta fecundities versions of the 
lottery model, specifically, in the subsequent sections.

%scaling factors
We now consider the scaling factors $q_{ij}$ for the lottery model. 
For the lottery model, for $i \neq j$, we have
\begin{align}
\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
\end{align}
Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$
in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
but we instead compute $q_{ij}$ through implicit differentiation:
\begin{align}
\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
\end{align}
Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so
\begin{align}
q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
\end{align}
This expression is made specific to the log-normal decundities and beta fecundities versions of the lottery
model in the subsequent sections. We alternatively also consider $q_{ij}=1$, for reasons explained in SI 
section \ref{SIsect:scaling_factors}.

%***DAN: Might need this text below, in the section on the log-normal fecundities lottery model, but not here
%Since we assumed log-normally distributed $B_i$, 
%$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
%to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
%and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).

%***DAN: We might need this text somewhere, but not sure where. Just saving it out of caution.
% For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
% $\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
% $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
% Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
% Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
% $\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
% can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
% is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
% and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
% marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
% $(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
% component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
% approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
% so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.

\subsection{Theory and plotting simplifications for the log-normal fecundities lottery model}\label{SIsect:TheoryForLotteryLogNorm}

%The first of the two growth rate decompositions
We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the 
log-normal fecundities lottery model. 

@@

E_i=B_i=\exp(\sigma b_i +\mu_i)

@@

C_{i \bs i} = B_j/\delta=\exp(\sigma b_j + \mu_j)/\delta

@@

\olr_{i \bs i} &= \E[\ln(1-\delta+\delta B_i/B_j)] \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]\label{eq:base_IGR}

@@

\olep_i^E &= \E[\ln(1-\delta+\delta\exp(\sigma b_i + \mu_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma b_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}

@@

where $b_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma b_j + \mu_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(-\sigma b_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 

@@

where $b_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(\sigma b_i +\mu_i)/\exp(\sigma b_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma (b_i - b_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}

@@

where $(b_i, b_j)$ is distributed as in SI section \ref{sect:noise}. Continuing,

@@

\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(\sigma b_i^\shp +\mu_i)/\exp(\sigma b_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i^\shp - b_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}

@@

where $b_i^\shp$ and $b_j^\shp$ are standard normally distributed and independent. Continuing,

@@

\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (b_i - b_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma (b_i^\shp - b_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}

@@

where $(b_i,b_j)$ is distributed as in SI section \ref{sect:noise}, and 
$b_i^\shp$ and $b_j^\shp$ are standard normal and independent. Continuing,

@@

\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_1}

@@

where $(b_i^\nat,b_j^\nat)$ have the same (standard normal) marginals and covariance as $(b_i,b_j)$ but are 
bivariate normal. Note, a choice has been made here: we could also have chosen $(b_i^\nat,b_j^\nat)$
to have the same standard normal marginals as $(b_i,b_j)$, but with covariance chosen such that
the Pearson, Spearman, or Kendall correlation of $E_i^\nat = \exp(\sigma b_i^\nat +\mu_i)$
and $C_{i \bs i}^\nat = \exp(\sigma b_j^\nat +\mu_j)/\delta$ matches that of $E_i = \exp(\sigma b_i +\mu_i)$
and $C_{i \bs i} = \exp(\sigma b_j +\mu_j)$. But the choice we made seems 
the most natural for the log-normal fecundities lottery models (and also for the 
beta fecundities lottery model - see below). 
Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_1})

@@

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (b_i^\shp - b_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}

@@

This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ 
for the log-normal fecundities lottery model.

%The second of the two growth rate decompositions
